---
title: "Tutorial"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=F}
library(learnr)
library(DBI)

knitr::opts_chunk$set(echo = FALSE)

orders <- read.delim('https://github.com/nchelaru/SQL-learnr/raw/master/orders.txt')

orderlines <- read.delim('https://github.com/nchelaru/SQL-learnr/raw/master/orderlines.txt')

products <- read.delim('https://github.com/nchelaru/SQL-learnr/raw/master/products.txt')

customers <- read.delim('https://github.com/nchelaru/SQL-learnr/raw/master/customers.txt')

campaigns <- read.delim('https://github.com/nchelaru/SQL-learnr/raw/master/campaigns.txt')

# Create an ephemeral in-memory RSQLite database
con <- dbConnect(RSQLite::SQLite(), ":memory:")

dbWriteTable(con, "orders", orders, overwrite=T)

dbWriteTable(con, "orderlines", orderlines, overwrite=T)

dbWriteTable(con, "products", products, overwrite=T)

dbWriteTable(con, "customers", customers, overwrite=T)

dbWriteTable(con, "campaigns", campaigns, overwrite=T)

```


## Getting to know the database

### List all tables in the database

Write the R code required to add two plus two:

```{r list-tables, exercise=TRUE, exercise.diagnostics=F}
dbListTables(con)
```

### List fields in table

Now write a function that adds any two numbers and then call it:

```{r list-fields, exercise=TRUE, exercise.diagnostics=F}
dbListFields(con, "products")
```

## Select data

### Exercise with Hint

```{r simple-select, exercise=TRUE, exercise.diagnostics=F}
query <- "SELECT * 
          FROM orders 
          LIMIT 10"

dbGetQuery(con, query)
```

## Filter data
```{r filter, exercise=TRUE, exercise.diagnostics=F}
query <- "SELECT * 
          FROM orders 
          WHERE city='BALTIMORE'"

dbGetQuery(con, query)
```

### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz}
quiz(
  question("Which package contains functions for installing other R packages?",
    answer("base"),
    answer("tools"),
    answer("utils", correct = TRUE),
    answer("codetools")
  ),
  question("Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```

